---
title: "CAOakland Exploration"
author: "Amber Lee"
date: "1/31/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

###  Connecting to the SQL server

```{r message = FALSE}
library(RMySQL)  
con <- dbConnect(
  MySQL(), host = "traffic.st47s.com", user = "<insert user here>", 
  password = "<insert password here>", dbname = "<insert dbname here>")
```

Next a series of SQL queries can be sent to the database using the `DBI::dbGetQuery()` function.

```{r message = FALSE}

#Each query returns an R dataframe
class(DBI::dbGetQuery(con, "SHOW TABLES"))

#Multiple tables within our traffic database
DBI::dbGetQuery(con, "SHOW TABLES")

#'Explain' command describe ten variables (field names) in the a table
DBI::dbGetQuery(con, "EXPLAIN CAoakland")

#Begin looking at Oakland databse
CAoakland_eightobsv <- DBI::dbGetQuery(con, "SELECT * FROM CAoakland LIMIT 8")
CAoakland_eightobsv

```


### Playing around with CAoakland

Exploring demographics of motorists stopped: race, sex, age, reason for stop

```{r}

db <- dplyr::src_mysql(dbname = "traffic", host = "traffic.st47s.com", user = "student", 
          password="Sagehen47")

CAoak <- dplyr::tbl(db, "CAoakland")

head(CAoak)

CAoak
dim(CAoak)
count(CAoak)
```

```{r Race and sex}

ggplot(data = CAoak) +
  geom_bar(mapping = aes(x = subject_race, fill = subject_sex)) +
  labs(title = "Subject race and sex counts")

```

* Note that males get stopped more in every single race category.

* Males are disproportionately stopped for API, hispanic, black, and "other" races.

* How does the distribution of traffic stop counts per race relate to the distribution of race in Oakland?

```{r Reason for stop}

#See the most frequent reason for stops
dbGetQuery(con, 
  "SELECT reason_for_stop, COUNT(reason_for_stop) AS 'number_reason_for_stop'
  FROM CAoakland
  GROUP BY(reason_for_stop)
  ORDER BY `number_reason_for_stop` ASC
  LIMIT 10")

#All reasons for stops
dbGetQuery(con, 
  "SELECT reason_for_stop, COUNT(reason_for_stop) AS 'number_reason_for_stop'
  FROM CAoakland
  GROUP BY(reason_for_stop)
  ORDER BY `number_reason_for_stop` ASC")

#Limit to the most frequent
CAoak_reason <- dbGetQuery(con, 
  "SELECT reason_for_stop, COUNT(reason_for_stop) AS 'number_reason_for_stop'
  FROM CAoakland
  GROUP BY(reason_for_stop)
  ORDER BY `number_reason_for_stop` DESC
  LIMIT 5")

CAoak_reason

#Total number of observations in the top 5 reason for stops
sum(CAoak_reason$number_reason_for_stop)

ggplot(data = CAoak_reason) +
  geom_bar(mapping = aes(x = reason_for_stop, y = number_reason_for_stop), stat = "identity") +
  labs(title = "Reason for stop")

```

* Traffic violation seems to be the most frequent reason for stop, but this is not counting all the duplicates!

* Would this be helpful to know? To clean the reason_for_stop?

```{r Temporality of stops}

#See the number of stops per day
CAoak_day_NA <- dbGetQuery(con, 
  "SELECT date, COUNT(date) AS 'number_stops_day'
  FROM CAoakland
  GROUP BY(date)
  ORDER BY `number_stops_day` ASC")

#Remove the NA column
CAoak_day <- CAoak_day_NA[-c(1),] 

library(lubridate)

#Use lubridate to make nice_day variable, which is just date-type variable
CAoak_day <- CAoak_day %>%
  mutate(nice_day = ymd(date))

#Plot number of stops throughout time
ggplot(data = CAoak_day, mapping = aes(x = nice_day, y = number_stops_day)) +
  geom_point() +
  geom_smooth() +
  labs(title = "Stop per day")
  
#Make year variable for facet wrap
CAoak_day <- CAoak_day %>%
  mutate(nice_year = year(nice_day))

#Plot number of stops throughout time, but per year
ggplot(data = dplyr::filter(CAoak_day, nice_year == 2013), mapping = aes(x = nice_day, y = number_stops_day)) +
  geom_point() +
  geom_smooth()

ggplot(data = dplyr::filter(CAoak_day, nice_year == 2014), mapping = aes(x = nice_day, y = number_stops_day)) +
  geom_point() +
  geom_smooth()

ggplot(data = dplyr::filter(CAoak_day, nice_year == 2015), mapping = aes(x = nice_day, y = number_stops_day)) +
  geom_point() +
  geom_smooth()

ggplot(data = dplyr::filter(CAoak_day, nice_year == 2016), mapping = aes(x = nice_day, y = number_stops_day)) +
  geom_point() +
  geom_smooth()

ggplot(data = dplyr::filter(CAoak_day, nice_year == 2017), mapping = aes(x = nice_day, y = number_stops_day)) +
  geom_point() +
  geom_smooth()

# I wasn't able to write a helpful function ):
# plot_years_CAoak <- function(start_year, end_year) { for (i_year in seq(start_year, end_year, 1)) 
#   {
# ggplot(data = dplyr::filter(CAoak_day, nice_year == i_year), mapping = aes(x = nice_day, y = number_stops_day)) +
#   geom_point() +
#   geom_smooth()
# }
# }

plot_years_CAoak(2013, 2017)

```

* Why is there a discontinuity in the data? in the last month of 2015?

* What is the pattern (or lack thereof) among each year?

* Note the abrupt jump from October 2014 to January 2015


```{r Age}

#Average age

dbGetQuery(con, 
  "SELECT subject_race, AVG(subject_age) AS 'ave age' 
  FROM CAoakland
  GROUP BY subject_race
  ORDER BY `ave age`")

```