---
title: "String Manipulation for Reason for Stop (CAoakland Exploration)"
author: "Amber Lee"
date: "1/31/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

###  Connecting to the SQL server

```{r message = FALSE}
library(RMySQL)  
con <- dbConnect(
  MySQL(), host = "traffic.st47s.com", user = "<insert user here>", 
  password = "<insert password here>", dbname = "<insert dbname here>")
```

Next a series of SQL queries can be sent to the database using the `DBI::dbGetQuery()` function.

```{r message = FALSE}

#Each query returns an R dataframe
class(DBI::dbGetQuery(con, "SHOW TABLES"))


#Multiple tables within our traffic database
DBI::dbGetQuery(con, "SHOW TABLES")

DBI::dbGetQuery(con, "EXPLAIN CAoakland")

#Begin looking at Oakland databse
CAoakland_eightobsv <- DBI::dbGetQuery(con, "SELECT * FROM CAoakland LIMIT 8")
CAoakland_eightobsv

# Random sample method, using rand()
# https://stackoverflow.com/questions/249301/simple-random-samples-from-a-sql-database
sample_CAoakland <- DBI::dbGetQuery(con, "SELECT * FROM CAoakland 
                                    WHERE rand() <=.05")
sample_CAoakland

```


### Playing around with CAoakland

Practice SQL commands to analyze "reason for stop"

```{r}

db <- dplyr::src_mysql(dbname = "traffic", host = "traffic.st47s.com", user = "student", 
          password="Sagehen47")

CAoak <- DBI::dbGetQuery(con, "SELECT * FROM CAoakland")

head(CAoak)

CAoak
dim(CAoak)
count(CAoak)
```

```{r Reason for stop}

#See the most frequent reason for stops
dbGetQuery(con, 
  "SELECT reason_for_stop, COUNT(reason_for_stop) AS 'number_reason_for_stop'
  FROM CAoakland
  GROUP BY(reason_for_stop)
  ORDER BY `number_reason_for_stop` ASC
  LIMIT 10")

#All reasons for stops
CAoak_reason <- dbGetQuery(con, 
  "SELECT reason_for_stop, COUNT(reason_for_stop) AS 'number_reason_for_stop'
  FROM CAoakland
  GROUP BY(reason_for_stop)
  ORDER BY `number_reason_for_stop` ASC")

#Limit to the most frequent
CAoak_reason_short <- dbGetQuery(con, 
  "SELECT reason_for_stop, COUNT(reason_for_stop) AS 'number_reason_for_stop'
  FROM CAoakland
  GROUP BY(reason_for_stop)
  ORDER BY `number_reason_for_stop` DESC
  LIMIT 5")

CAoak_reason

#Total number of observations in the top 5 reason for stops
sum(CAoak_reason$number_reason_for_stop)

ggplot(data = CAoak_reason_short) +
  geom_bar(mapping = aes(x = reason_for_stop, y = number_reason_for_stop), stat = "identity") +
  labs(title = "Top 5 reasons for stop")

```


```{r reason_for_stop column SQL demand}

#Column with all reason_for_stop (length 133407)
reason_for_stop_column <- dbGetQuery(con, 
  "SELECT reason_for_stop FROM CAoakland")

reason_for_stop_column

CAoak_reason


```

### Attempt to clean up strings using string manipulation and regular expressions

Decided not to continue because stops with weirdly labelled "reason for stop" was rare

```{r Get rid of comma and isolate duplicates}

# Remove comma
reason_for_stop_column <- mutate(reason_for_stop_column, no_comma = str_replace_all(reason_for_stop, ",", ""))

# Check to see that comma entries are gone
reason_for_stop_column %>%
  filter(str_detect(no_comma, ","))

# Return TRUE if the | (pipe character) is found in the reason_for_stop entry
# Pipe character | designates that there are multiple and/or duplicate reasons for stop
str_detect(reason_for_stop_column[, 1], "\\|")

# Table of reason_for_stop entries that have the pipe character
reason_duplicates <- reason_for_stop_column %>%
  filter(str_detect(reason_for_stop, "\\|"))

reason_duplicates

# Function to remove duplicates from StackOverflow
# https://stackoverflow.com/questions/20283624/removing-duplicate-words-in-a-string-in-r

rem_dup.one <- function(x){
  paste(unique(trimws(unlist(strsplit(x,split="\\|",fixed=F,perl=T)))),collapse = " ")
}

# Testing individual strings will work
rem_dup.one("Reasonable Suspicion|Reasonable Suspicion|Reasonable Suspicion|Probable Cause")
rem_dup.one("Reasonable Suspicion|Probable Cause|Probable Cause")

# But upon using mutate, function will bug.
reason_duplicates <- reason_duplicates %>% 
  mutate(no_duplicates = rem_dup.one(no_comma))

reason_duplicates

```

I can't find a systematic way to get rid of the pipes. So I'm going to:

Directly normalize the data

* create list of official reasons

* go through each entry in reason_for_stop_column SPLIT by \\| pipe character

* add onto reason_for_stop_column columns that are official reasons

* if the split entry has entires that are in official reason_i, mark 1 in the official reasons column or 0

### Not string-related stuff

I want to try EDA with other people's code! Because I don't know how to debug.

# Will's

```{r}
Coordinates <- DBI::dbGetQuery(con,
  "SELECT lng, lat, subject_race FROM CAoakland WHERE rand() <=.02")

coor <- DBI::dbGetQuery(con,
  "SELECT lng, lat, subject_race FROM CAoakland")

# Note that lng and lat are of type double 
  
race_plot <- ggplot(Coordinates, aes(x = as.numeric(lng), y = as.numeric(lat), xaxt = 'n', yaxt = 'n')) + 
  geom_point(aes(color = subject_race), position = "jitter", size = .25) + 
  xlab("Longitude") + 
  ylab("Latitude") + 
  coord_quickmap()
race_plot

# Facet wrap race_plot
ggplot(Coordinates, aes(x = as.numeric(lng), y = as.numeric(lat), xaxt = 'n', yaxt = 'n')) + 
  geom_point(aes(color = subject_race), position = "jitter", size = .25) + 
  xlab("Longitude") + 
  ylab("Latitude") + 
  coord_quickmap() +
  facet_wrap( ~ subject_race, nrow = 2)

```


```{r}

library(sf)
library(mapview)

# Mapview basics
# https://r-spatial.github.io/mapview/articles/articles/mapview_02-advanced.html
mapview(breweries)
View(breweries)
View(franconia)
typeof(breweries$address)


# Create empty geometries list to fill in later
# Source: https://stackoverflow.com/questions/46385301/replace-geometries-from-a-list-in-sf
nrows = nrow(CAoak)
geometry = st_sfc(lapply(1:nrows, function(x) st_geometrycollection()))
df <- st_sf(id = 1:nrows, geometry = geometry)
df

for (i in 1:nrows){
  geo_singletrafficstop = st_point(c(CAoak$lat[i], CAoak$lng[i]))
  df$geometry[i] = geo_singletrafficstop
}

ef <- cbind(df, CAoak$subject_race)

#saving maps -- these are ugly/ I want to zoom around
m <- mapview(ef, zcol = "sample_CAoakland.subject_race", legend = TRUE)


```
