---
title: "Summary of Weeks 1-8"
author: "Amber Lee"
date: "4/27/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(RMySQL)
library(suncalc)
library(lutz)
library(lubridate)

# Connect to Oakland dataset
con <- dbConnect(
  MySQL(), host = "traffic.st47s.com", user = "<insert user here>", 
  password = "<insert password here>", dbname = "<insert dbname here>")

CAoak <- DBI::dbGetQuery(con, "SELECT * FROM CAoakland")

# Lubridate
CAoak <- CAoak %>%
  filter(!is.na(date)) %>%
  mutate(nice_date = ymd(date), 
         nice_year = year(nice_date), 
         nice_month = month(nice_date), 
         nice_day = day(nice_date), 
         nice_time = hms(time),
         nice_day_of_year = yday(date),
         
         #for sunset sunrise: 
         posix_date_time = as.POSIXct(paste(nice_date, time), tz = "America/Los_Angeles", format = "%Y-%m-%d %H:%M:%OS"))

```

```{r sunrise/sunset set up}

OaklandTZ <- lutz::tz_lookup_coords(37.75241810000001, -122.18087990000001, warn = F)
# SanAntonioTZ <- lutz::tz_lookup_coords(29.4241, -98.4936, warn = F)

temp_data <- data.frame(date = CAoak$nice_date, lat = CAoak$lat, lon = CAoak$lng)

getSunlightTimes(data = temp_data, keep = c("sunrise", "sunset"), tz = "America/Los_Angeles") %>% head()

oursunriseset <- function(latitude, longitude, date, direction = c("sunrise", "sunset")) {
  date.lat.long <- data.frame(date = date, lat = latitude, lon = longitude)
  if(direction == "sunrise"){
    getSunlightTimes(data = date.lat.long, keep=direction, tz = OaklandTZ)$sunrise }else{
      getSunlightTimes(data = date.lat.long, keep=direction, tz = OaklandTZ)$sunset } }

# add light variable 
CAoak <- CAoak %>%
  
  # use oursunriseset function to return posixct format sunrise and sunset times
  mutate(sunrise = oursunriseset(lat, lng, nice_date, direction = "sunrise"),
         sunset = oursunriseset(lat, lng, nice_date, direction = "sunset")) %>%

  mutate(light = ifelse(posix_date_time > sunrise & posix_date_time < sunset, "day", "night")) %>%
  
  # about 100 NA's to filter out

  filter(!is.na(light))

```

# Race-related Visualizations

## To be included in the final report

```{r initial visualizations race}

# Race and sex

ggplot(data = CAoak) +
  geom_bar(mapping = aes(x = subject_race, fill = subject_sex))
ggsave("race and sex.png")

# Race and age

# 102,724 (77%) of the observations *do not* have subject_age recorded
# CAoak %>%
#   group_by(subject_age) %>%
#   summarize(count = n())

## Age is only recorded for the year 2017

CAoak %>%
  filter(!is.na(subject_age)) %>%
  group_by(nice_date, subject_race) %>%
  summarize(count = n()) %>%
  ggplot(aes(x = nice_date, y = count, color = subject_race)) +
  geom_point()

# average age of stop for Black and Latinx is lower than for average age of white and API

dbGetQuery(con, 
  "SELECT subject_race, AVG(subject_age) AS 'ave age' 
  FROM CAoakland
  GROUP BY subject_race
  ORDER BY `ave age`")

CAoak %>%
  filter(subject_age < 50 & subject_age > 18) %>%
  filter(subject_race != "asian/pacific islander" & subject_race != "other") %>%
  ggplot +
  geom_bar(mapping = aes(x = subject_age, fill = subject_race), position = "dodge")
ggsave("age of motorist by race, BWH.png")

CAoak %>%
  filter(!is.na(subject_age), subject_age >= 20 & subject_age < 60) %>%
  mutate(age_group = case_when(subject_age >= 20 & subject_age < 25 ~ "20 to 24",
                               subject_age >= 25 & subject_age < 30 ~ "25 to 30",
                               subject_age >= 30 & subject_age < 35 ~ "30 to 34",
                               subject_age >= 35 & subject_age < 40 ~ "35 to 39",
                               subject_age >= 40 & subject_age < 45 ~ "40 to 44",
                               subject_age >= 45 & subject_age < 50 ~ "45 to 49",
                               subject_age >= 50 & subject_age < 55 ~ "50 to 54",
                               subject_age >= 55 & subject_age < 60 ~ "55 to 60")) %>%
  ggplot(aes(x = age_group, fill = subject_race)) +
  geom_bar(position = "dodge")
ggsave("age of motorist by race, all races and binned ages.png")

```

## NOT to be included in the final report

```{r}

# outcome and race
CAoak %>%
  ggplot()+
  geom_bar(mapping = aes(x = subject_race, fill = outcome), position = "dodge")
ggsave("outcome and race.png")

```


## Incorporating Day/Night Variable in race analysis 

```{r}

# Day/night

# see how drivers are stopped by race and light in absolute counts
ggplot(data = CAoak) +
  geom_bar(mapping = aes(x = subject_race, fill = light))
ggsave("night and day stop counts by race.png")

# percentage of each racial group stopped during the day
CAoak %>%
  group_by(subject_race, light) %>%
  summarize(count = n()) %>%
  spread(key = light, value = count) %>%
  mutate(total_stops = sum(day, night),
         perc_light = day / total_stops) %>%
  ggplot(aes(x = subject_race, y = perc_light)) +
  geom_bar(stat = "identity")
ggsave("night and day stop percents by race.png")

# how are outcomes of stops affected by race and time of day?
  
CAoak %>%
  ggplot(aes(x = subject_race, fill = outcome, color = outcome, alpha = light)) +
  geom_bar(position="dodge") + 
  scale_alpha_manual(values=c(.2, .9))
ggsave("outcome and race and light.png")

CAoak %>%
  ggplot() +
  geom_bar(aes(x = search_conducted, fill = subject_race)) +
  facet_wrap(~ outcome)

```

### TODO: MAKE outcome, race, and light/day 

# Outcomes
find the difference between # of drivers with search conducted - # of drivers with search conducted and arrest made, for each age and race

## Compare ALL stops with stop AND search

```{r}

stop_searches <- CAoak %>%
  
  # count stops by search or no search
  group_by(subject_race, subject_age, search_conducted) %>%
  summarize(count = n())

# stops regardless of search was conducted
stops_and_searches <- CAoak %>%
  ungroup() %>%
  group_by(subject_race, subject_age) %>%
  summarize(count = n()) %>%
  select(subject_race, subject_age, count) %>%
  rename(stopsearch_count = count)

# stops that DID result in searches
searches_only <- stop_searches %>%
  ungroup() %>%
  filter(search_conducted == 1) %>%
  select(subject_race, subject_age, count) %>%
  rename(search_count = count)

# absolute numbers
stops_and_searches %>%
  full_join(searches_only, by = c("subject_race", "subject_age")) %>%
  
  # number of stops WITHOUT a search minus stops WITH a search
  mutate(difference_stop_search = stopsearch_count - search_count) %>%
  ggplot() + 
  geom_bar(aes(x = subject_age, y = difference_stop_search, fill = subject_race), stat = "identity") +
  facet_wrap(~ subject_race)

ggsave("all stop and search differences absolute.png")


#percentages
stops_and_searches %>%
  full_join(searches_only, by = c("subject_race", "subject_age")) %>%
  
   # number of stops WITHOUT a search minus stops WITH a search
  mutate(difference_stop_search = (stopsearch_count - search_count)/(stopsearch_count + search_count)) %>%
  ggplot() + 
  geom_bar(aes(x = subject_age, y = difference_stop_search, fill = subject_race), stat = "identity") +
  facet_wrap(~ subject_race)

ggsave("all stop and search differences percent.png")

```

## Compare stops AND NO search and stops AND search

```{r}

stop_searches <- CAoak %>%
  
  # count stops by search or no search
  group_by(subject_race, subject_age, search_conducted) %>%
  summarize(count = n())

# stops that DID NOT result in searches
stops_no_search <- stop_searches %>%
  ungroup() %>%
  filter(search_conducted == 0) %>%
  select(subject_race, subject_age, count) %>%
  rename(stop_count = count)

# stops that DID result in searches
# searches_only 

# absolute numbers
stops_no_search %>%
  full_join(searches_only, by = c("subject_race", "subject_age")) %>%
  
  # number of stops WITHOUT a search minus stops WITH a search
  mutate(difference_stop_search = stop_count - search_count) %>%
  ggplot() + 
  geom_bar(aes(x = subject_age, y = difference_stop_search, fill = subject_race), stat = "identity") +
  facet_wrap(~ subject_race)

ggsave("stop and search differences absolute.png")


#percentages
stops_no_search %>%
  full_join(searches_only, by = c("subject_race", "subject_age")) %>%
  
   # number of stops WITHOUT a search minus stops WITH a search
  mutate(difference_stop_search = (stop_count - search_count)/(search_count+stop_count)) %>%
  ggplot() + 
  geom_bar(aes(x = subject_age, y = difference_stop_search, fill = subject_race), stat = "identity") +
  facet_wrap(~ subject_race)

ggsave("stop and search differences percent.png")

```

## Compare ALL searches with search AND arrest

```{r}

search_arrests <- CAoak %>%
  
  # filter for drivers who are stopped, b/w/h
  #filter(search_conducted == 1, subject_race != "asian/pacific islander", subject_race != "other", subject_age > 14, subject_age < 70) %>%
  
  # count number of stops per age 
  group_by(subject_race, subject_age, search_conducted, arrest_made) %>%
  summarize(count = n())

# stops that resulted in searches
# searches_only

# stops that resulted in search and arrests
search_and_arrests <- search_arrests %>%
  ungroup() %>%
  filter(search_conducted == 1, arrest_made == 1) %>%
  select(subject_race, subject_age, count) %>%
  rename(arrest_count = count)

# absolute numbers
searches_only %>%
  full_join(search_and_arrests, by = c("subject_race", "subject_age")) %>%
  
  # number of stops WITHOUT a search minus stops WITH a search
  mutate(difference_stop_search = search_count - arrest_count) %>%
  ggplot() + 
  geom_bar(aes(x = subject_age, y = difference_stop_search, fill = subject_race), stat = "identity") +
  facet_wrap(~ subject_race)

# ggsave("all stop and search differences absolute.png")


#percentages
searches_only %>%
  full_join(search_and_arrests, by = c("subject_race", "subject_age")) %>%
  
   # number of stops WITHOUT a search minus stops WITH a search
  mutate(difference_stop_search = (search_count - arrest_count)/(search_count + arrest_count)) %>%
  ggplot() + 
  geom_bar(aes(x = subject_age, y = difference_stop_search, fill = subject_race), stat = "identity") +
  facet_wrap(~ subject_race)

# ggsave("all stop and search differences percent.png")

```

## Compare search AND NO arrest with search AND arrest

```{r}

# looking at stops that led to at least a search
# search_arrests

# count searches that led to no arrest
searches_no_arrest <- search_arrests %>%
  filter(search_conducted == 1, arrest_made == 0) %>%
  select(subject_race, subject_age, count) %>%
  rename(search_count = count)

# count searches that led to arrests
# search_and_arrests

# absolute numbers
searches_no_arrest %>%
  full_join(search_and_arrests, by = c("subject_race", "subject_age")) %>%
  mutate(difference_search_arrest = search_count - arrest_count) %>%
  ggplot() + 
  geom_bar(aes(x = subject_age, y = difference_search_arrest, fill = subject_race), stat = "identity") +
  facet_wrap(~ subject_race)

ggsave("search and arrest differences absolute.png")

#percentages
searches_no_arrest %>%
  full_join(search_and_arrests, by = c("subject_race", "subject_age")) %>%
  mutate(difference_search_arrest = (search_count - arrest_count)/(search_count+arrest_count)) %>%
  ggplot() + 
  geom_bar(aes(x = subject_age, y = difference_search_arrest, fill = subject_race), stat = "identity") +
  facet_wrap(~ subject_race)

ggsave("search and arrest differences percent.png")

```

Searches AND no arrest minus searches AND arrest.
Willy nilly searches?

When difference_search_arrest is positive, then the number of high-discretionary searches (meaning, the number of searches that occured *without* resulting in an arrest, so police had high discretion in pulling those drivers over) is high. We see that the percent of high-discretionary stops over all discretionary stops is almost entirely positive for black drivers under 50 and hispanic drivers under 30. 

# Reason for stop investigation

```{r reason for stop}

# Most frequent reason for stops
frequent_reasons_CAoak <- dbGetQuery(con, 
  "SELECT reason_for_stop, COUNT(reason_for_stop) AS 'number_reason_for_stop'
  FROM CAoakland
  GROUP BY(reason_for_stop)
  ORDER BY `number_reason_for_stop` DESC
  LIMIT 5")

frequent_reasons_CAoak

# Least frequent reason for stops (messy data)

dbGetQuery(con, 
  "SELECT reason_for_stop, COUNT(reason_for_stop) AS 'number_reason_for_stop'
  FROM CAoakland
  GROUP BY(reason_for_stop)
  ORDER BY `number_reason_for_stop` ASC
  LIMIT 5")

# Plot absolute and relative number of stops for frequent reasons

ggplot(data = frequent_reasons_CAoak) +
  geom_bar(mapping = aes(x = reason_for_stop, y = number_reason_for_stop), stat = "identity")

frequent_reasons_CAoak %>%
  mutate(perc = number_reason_for_stop/sum(frequent_reasons_CAoak$number_reason_for_stop)) %>%
  ggplot(aes(x = reason_for_stop, y = perc)) +
  geom_bar(stat = "identity")

# Relative number of cleanly coded reason for stop
sum(frequent_reasons_CAoak$number_reason_for_stop)/dim(CAoak)[1]


```

Most of the 133,407 traffic stops in Oakland have a cleanly-coded reason for stop. The most frequently cited reason is traffic violation (70+%).

# Maps

```{r}

coor <- DBI::dbGetQuery(con, "SELECT lng, lat, subject_race FROM CAoakland")

# Note that lng and lat are of type double 
  
ggplot(coor, aes(x = as.numeric(lng), y = as.numeric(lat), xaxt = 'n', yaxt = 'n')) + 
  geom_point(aes(color = subject_race), position = "jitter", size = .25) + 
  xlab("Longitude") + 
  ylab("Latitude") + 
  coord_quickmap()

# Facet wrap race_plot
ggplot(coor, aes(x = as.numeric(lng), y = as.numeric(lat), xaxt = 'n', yaxt = 'n')) + 
  geom_point(aes(color = subject_race), position = "jitter", size = .25) + 
  xlab("Longitude") + 
  ylab("Latitude") + 
  coord_quickmap() +
  facet_wrap( ~ subject_race, nrow = 2)
```

#To do: add the empirical search probabilities and make it pretty

```{r}
# create list of all dataset names
all_dataset_names <-as.list(DBI::dbGetQuery(con, "SHOW TABLES"))$Tables_in_traffic
# all_dataset_names

variables_of_interest <- c("subject_age", "subject_race", "subject_sex", "date", "search_conducted")

relevant_datasets <- function(all_dataset_names, variables_of_interest){
  
  # create empty vector
  datasets_of_interest <- c()
    
  for(city in all_dataset_names){
    
    # cancenate SQL query string
    command <- paste("EXPLAIN", city, sep = " ")
    field_vector <- unlist(as.list(DBI::dbGetQuery(con, command))$Field, 
                           use.names = FALSE)
    
    # able to check if subject_age lives as a different variable name
    # print(field_vector)
    
    # of_interest_book is TRUE iff field_vector contains all the variables of interest
    of_interest_bool <- setequal(intersect(field_vector, variables_of_interest),
                                 variables_of_interest)
    
    # add dataset name to vector if of_interest_bool
    if(of_interest_bool){
      datasets_of_interest <- c(datasets_of_interest, city)
    }
  }
  
  return(datasets_of_interest)
}

datasets_of_interest <- relevant_datasets(all_dataset_names, variables_of_interest)
datasets_of_interest
```

a) query each dataset of interest (27 in total), just the variables of interest (search conducted, subject race, etc.). (optional: maybe randomly sample 10% from each dataset). for loop through datasets of interest and download. -- DONE!
```{r}

datasets_of_interest
truncate_datasets = datasets_of_interest[2:6]

for(city in datasets_of_interest){
  
  # cancenate SQL query string
  # save only .1% of dataset bc my computer is tired
  command <- paste("SELECT subject_age, subject_race, subject_sex, date, search_conducted FROM", city, "WHERE rand() <= .01", sep = " ")
  
  # save dataset with unique names 
  assign(paste(city), DBI::dbGetQuery(con, command))
  # make a global variable in assign thru envir = .GlobalEnv
  
  }

```

go thru datasets
1. download total stop counts by race and total search counts by race
2. join the two datasets
3. calculate (rolling) avg. probability of being searched
4. create 4 column df
5. cbind the df from #4 to the original initialized df
```{r}

# function whose input is a list of datasets
# return one giant dataset

p_search_conducted <- function(dataset_name){

  # first, check the type of search_conducted using EXPLAIN
  
  dataset_str <- paste(dataset_name)
  
  explain_command_str <- paste("EXPLAIN", dataset_name, sep = " ")
  
  explain_df <- dbGetQuery(con, explain_command_str) %>%
    filter(Field == "search_conducted") %>%
    mutate(dataset = dataset_str) %>%
    select(Field, Type, dataset)
  
  # note that explain_df[1, 2] is the entry that has the type of search_conducted
  
  # second, create SQL search strings based on type of search_conducted
  
  if (explain_df[1, 2] == "varchar(50)") {
    
      search_numerator_sql <- paste("SELECT subject_age, subject_race, 
                                COUNT(*) as 'search_counts' FROM",
                                dataset_str, 
                                "WHERE (subject_race='black' 
                                OR subject_race = 'white') 
                                AND search_conducted = 'TRUE' 
                                AND subject_age > 0 
                                GROUP BY subject_age, subject_race", sep = " ")
    
  } else if (explain_df[1, 2] == "double") {
    
      search_numerator_sql <- paste("SELECT subject_age, subject_race, 
                                COUNT(*) as 'search_counts' FROM",
                                dataset_str, 
                                "WHERE (subject_race='black' 
                                OR subject_race = 'white') 
                                AND search_conducted = '1' 
                                AND subject_age > 0 
                                GROUP BY subject_age, subject_race", sep = " ")
      
  }
  
  stops_denominator_sql <- paste("SELECT subject_age, subject_race, 
                                 COUNT(*) as 'total_stop_counts' FROM", 
                                 dataset_str, 
                                 "WHERE (subject_race='black' 
                                 OR subject_race = 'white') 
                                 AND subject_age > 0 
                                 GROUP BY subject_age, subject_race", sep = " ")
  
  # third, calculate % search_conducted per age in df thru query
  
  search_numerator <- dbGetQuery(con, search_numerator_sql)
  
  stops_denominator <- dbGetQuery(con, stops_denominator_sql)
  
  # fourth, combine results into one df 
  
  search_probability <- search_numerator %>%
    right_join(stops_denominator, by = c("subject_race", "subject_age")) %>%
    replace_na(list(search_counts = 0)) %>%
    mutate(search_percent = search_counts / total_stop_counts,
           
           # create column for dataset name
           dataset = dataset_str,
           subject_age = as.numeric(subject_age))
  
  return(search_probability)

}

search_probs_list <- lapply(datasets_of_interest, p_search_conducted)

search_probs_list

combined_search_probs_list <- bind_rows(search_probs_list, .id = "column_label")

combined_search_probs_list %>%
  filter(dataset != "PApittsburgh") %>%
  ggplot() +
  geom_point(mapping = aes(x = subject_age, y = search_percent, color = subject_race), alpha = .4) +
  facet_wrap(~ dataset) +
  scale_y_continuous(limits = c(0, .4))
```
