---
title: "Night percent, logistic regression"
author: "Amber Lee"
date: "3/7/2020"
output: pdf_document
---

# Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shiny)
library(suncalc)
library(lubridate)
library(tidyverse)
library(RMySQL)  
library(caret)
library(rpart)

con <- dbConnect(
  MySQL(), host = "traffic.st47s.com", user = "<insert user here>", 
  password = "<insert password here>", dbname = "<insert dbname here>")

sample_CAoakland <- DBI::dbGetQuery(con, "SELECT * FROM CAoakland 
                                    WHERE rand() <=.05")

CAoak <- DBI::dbGetQuery(con, "SELECT * FROM CAoakland")

```

# Night percent - day percent

```{r prettily implementing the percent night - percent day in oakland}

# 1. fix the dates and lat/lng types. Check correct timezones for tz

CAoak <- CAoak %>%
  
  # optional: i have to filter out my NA's for date for POSIX to work
  filter(str_detect(date, "NA", negate = TRUE)) %>%
  
  mutate(nice_date = ymd(date), 
         nice_year = year(nice_date), 
         nice_month = month(nice_date), 
         nice_day = day(nice_date), 
         nice_time = hms(time),
         nice_day_of_year = yday(date),
         #for sunset sunrise: 
         posix_date_time = as.POSIXct(paste(nice_date, time), tz = "America/Chicago", format = "%Y-%m-%d %H:%M:%OS")) %>%
  mutate(lat_num = as.numeric(lat),
         lng_num = as.numeric(lng)) 

# 2. use sunrise/sunset function, again heeding the tz

oursunriseset <- function(latitude, longitude, date, direction = c("sunrise", "sunset")) {
  date.lat.long <- data.frame(date = date, lat = latitude, lon = longitude)
  if(direction == "sunrise"){
    getSunlightTimes(data = date.lat.long, keep=direction, tz = "America/Los_Angeles")$sunrise }else{
      getSunlightTimes(data = date.lat.long, keep=direction, tz = "America/Los_Angeles")$sunset } }

# 3. create variable for light (day/night)

CAoak <- CAoak %>%
  
  # use oursunriseset function to return posixct format sunrise and sunset times
  mutate(sunrise = oursunriseset(lat_num, lng_num, nice_date, direction = "sunrise"),
         sunset = oursunriseset(lat_num, lng_num, nice_date, direction = "sunset")) %>%

  # night and day!!
  mutate(light = ifelse(posix_date_time > sunrise & posix_date_time < sunset, "day", "night")) 

# 4a. count the number of ALL DRIVERS and BLACK DRIVERS stopped during day and night.
# 4b. calculate the percentage of black/all for day AND black/all for night

CAoakcheckpoint <- CAoak %>%
  
  # filter out the NA's for light variable
  filter(light == "day" | light == "night") %>%
  
  # group by month, year, and light
  group_by(nice_month, nice_year, light) %>%
  
  # count number of drivers stopped per month during night/day
  summarise(all_drivers_stopped = n(), black_drivers_stopped = sum(subject_race == "black")) %>%
  
  # find percent of black/all drivers stopped for day and night
  mutate(stops_black_percent = black_drivers_stopped/all_drivers_stopped) %>%
  
  #create arbitrary lubridate (first day of each month) for each year-month pair
  mutate(month_year = ymd(paste(nice_year, nice_month, "1", sep = "-"))) 

# 5. use filter to create two seperate day and night dataframes (to be joined later)

CAoak_day_stopcounts <- CAoakcheckpoint %>% filter(light == "day")
CAoak_night_stopcounts <- CAoakcheckpoint %>% filter(light == "night")

# 6. join and use mutate to calculate percents day/night and percent differences

# join by month_year
# do keep: all_drivers_stopped, black_drivers_stopped, and stops_black_percent for both day, night
# 6 variables in total

CAoak_join_stopcounts <- inner_join(CAoak_day_stopcounts, CAoak_night_stopcounts, by = c("month_year", "nice_month", "nice_year")) %>%
  
  # rename columns for clarity (day/night)
  rename(day_all_drivers_stopped = all_drivers_stopped.x,
         night_all_drivers_stopped = all_drivers_stopped.y,
         day_black_drivers_stopped = black_drivers_stopped.x,
         night_black_drivers_stopped = black_drivers_stopped.y,
         day_stops_black_percent = stops_black_percent.x,
         night_stops_black_percent = stops_black_percent.y) %>%
  
  # calculate the difference! OBSERVE that it is night percent difference
  mutate(racial_percent_diff = night_stops_black_percent - day_stops_black_percent)

CAoak_join_stopcounts %>%
  
  ggplot(mapping = aes(x = month_year, y = racial_percent_diff))+
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0, color = "red") +
  labs(y = "% Black stopped night - % Black stopped day")

#ggsave("CAoak_daynightpercent_tidyversemethod.png")

```

# Questions: Have the day traffic stops and night traffic stop relative proportions stayed the same?

To answer this question, I build off of the already-cleaned CAoak_join_stopcounts

```{r percent of night stops and total stops}

CAoak_join_stopcounts <- CAoak_join_stopcounts %>%
  
  # Find all stop counts
  mutate(total_stop_count = day_all_drivers_stopped + night_all_drivers_stopped,
         
         # Find percentage of night-time stops
         nighttime_percent = night_all_drivers_stopped/total_stop_count)

CAoak_join_stopcounts %>%
  
  ggplot(mapping = aes(x = month_year, y = nighttime_percent)) +
  geom_point(color = "blue") +
  geom_line(color = "blue")

ggsave("CAoak_nightpercent.png")

CAoak_join_stopcounts %>%
  
  # Find all stop counts
  mutate(total_stop_count = day_all_drivers_stopped + night_all_drivers_stopped,
         
         # Find percentage of night-time stops
         nighttime_percent = night_all_drivers_stopped/total_stop_count) %>%
  
  ggplot(mapping = aes(x = month_year, y = nighttime_percent)) +
  geom_point(color = "blue") +
  geom_line(color = "blue") +
  
  # Overlay the racial_percent_diff from earlier chunk
  geom_point(mapping = aes(x = month_year, y = racial_percent_diff)) +
  geom_line(mapping = aes(x = month_year, y = racial_percent_diff)) +
  geom_hline(yintercept = 0, color = "red")

ggsave("CAoak_overlaydaynightpercent_tidyversemethod.png")
  

# number of stops per month

CAoak_join_stopcounts %>%
  
  ggplot(mapping = aes(x = month_year, y = total_stop_count)) +
  geom_point() +
  geom_line()

# Count the total number of stops: 41k. 41k out of 133k is about 30%, so using this data to model searches conducted for drivers stopped would be worthwhile (41k is big enough) 
CAoak %>%
  
  filter(search_conducted == "TRUE") %>%
  summarise(n())


```

# Interlude: light, subject race, outcome

```{r}
  
CAoak %>%
  
  #filter out NA's for readability
  filter(!is.na(light)) %>%
  ggplot(aes(x = subject_race, fill = outcome, alpha = light)) +
  geom_bar(position="dodge", colour="black") + coord_flip() + scale_alpha_manual(values=c(.2, .8))

CAoak %>%
  
  #filter out NA's for readability
  filter(!is.na(light)) %>%
  filter(search_conducted == "1") %>%
  ggplot(aes(x = subject_race, fill = outcome, alpha = light)) +
  geom_bar(position="dodge", colour="black") + coord_flip() + scale_alpha_manual(values=c(.2, .8))

CAoak %>%
  
  #filter out NA's for readability
  filter(!is.na(light)) %>%
  filter(search_conducted == "0") %>%
  ggplot(aes(x = subject_race, fill = outcome, alpha = light)) +
  geom_bar(position="dodge", colour="black") + coord_flip() + scale_alpha_manual(values=c(.2, .8))

CAoak %>%
  
  #filter out NA's for readability
  filter(!is.na(light)) %>%
  filter(search_conducted == "1") %>%
  ggplot(aes(x = subject_race, fill = outcome, alpha = light)) +
  geom_bar(position="dodge", colour="black") + coord_flip() + scale_alpha_manual(values=c(.2, .8))

ggsave("outcome, nightday, race progress.png")

```

perhaps can look at searches, then outcome :) 
searches precede citation, arrest. read more about the process of getting into the criminal justice system


```{r}

ggplot(data = CAoak) + 
   geom_bar(mapping = aes(x = as.numeric(subject_age), fill = subject_race)) +
  facet_wrap(~ subject_race)

CAoak %>%
  filter(search_conducted == "1") %>%
  ggplot() + 
  geom_bar(mapping = aes(x = as.numeric(subject_age), fill = subject_race)) +
  facet_wrap(~ subject_race)

CAoak %>%
  filter(search_conducted == "1", arrest_made == "1") %>%
  ggplot() + 
  geom_bar(mapping = aes(x = as.numeric(subject_age), fill = subject_race)) +
  facet_wrap(~ subject_race)

# Question: do these distributions reflect census data?


```

### Logistic Regression

```{r logistic regression all searches}

logreg_oak1 <- CAoak %>%
  
  #only 30k out of 133k of my data records subject age
  filter(subject_age != "NA") %>%
  
  #use case_when to recode character variables to binary levels 
  mutate(
         # assigned day = 1
         light_binary = case_when(light == "day" ~ 1,
                                  light == "night" ~ 0),
         subject_age = as.numeric(subject_age)) %>%
  select(subject_age, search_conducted, search_conducted, light, light_binary, subject_race, arrest_made, lat_num, lng_num, subject_sex)

all_output1 <- glm(formula = search_conducted ~ subject_age*subject_race + factor(light_binary), data = logreg_oak1, family = binomial)

summary(all_output1)

```
concerns:
1) this is looking at all stops vs. all stops + searches. may want to look at all stops vs. searches THAT DIDN'T RESULT IN AN ARREST
2) may want to bin ages into rough age groups

```{r logistic regression discretionary searches}

# count the number of searches conducted that did and didn't result in arrests
CAoak %>%
  select(search_conducted, arrest_made) %>%
  group_by(search_conducted, arrest_made) %>%
  summarise(n())

# out of 41,156 searches conducted, 15870 resulted in arrests made. that is 40%

# conduct logistic regression looking at search conducted but arrest not made

logreg_oak2 <- logreg_oak1 %>%
  filter(arrest_made == "0")

all_output2 <- glm(formula = search_conducted ~ subject_age*subject_race + factor(light_binary), data = logreg_oak2, family = binomial)

summary(all_output2)

```

* note the coefficients that become statistically significant when looking only at discretionary searches:

** all searches **
> Coefficients:
                                   Estimate Std. Error z value Pr(>|z|)    
subject_age                      -0.0359989  0.0052561  -6.849 7.44e-12 ***
subject_raceblack                 0.3525380  0.1859483   1.896 0.057974 .  
subject_racewhite                 0.0611644  0.2340115   0.261 0.793805    
factor(light_binary)1            -0.0101902  0.0257883  -0.395 0.692734    
subject_age:subject_raceblack     0.0184709  0.0053817   3.432 0.000599 ***
subject_age:subject_raceother    -0.0256453  0.0103505  -2.478 0.013224 *  
subject_age:subject_racewhite    -0.0004428  0.0065765  -0.067 0.946314  

** discretionary searches only **
> Coefficients:
                                   Estimate Std. Error z value Pr(>|z|)    
subject_age                      -0.0516224  0.0070314  -7.342 2.11e-13 ***
subject_raceblack                 0.1766233  0.2311830   0.764 0.444869    
subject_racewhite                -0.6250167  0.2971249  -2.104 0.035418 *  
factor(light_binary)1             0.1618909  0.0305709   5.296 1.19e-07 ***
subject_age:subject_raceblack     0.0253226  0.0071697   3.532 0.000413 ***
subject_age:subject_raceother    -0.0145676  0.0135071  -1.079 0.280805    
subject_age:subject_racewhite     0.0147832  0.0087541   1.689 0.091275 .  

* the magnitude of subject_age coefficient increases for discretionary searches (-.03 to -.05, more significant)

* magnitude of subject_racewhite coefficient goes from .06 insignificant to -.625 statistically significant when limiting to discretionary searches

* factor(light_binary) becomes positive .16 and statistically significant when limiting to discretionary searches. day = 1 and night = 0, so how to interpret the +.16 coefficient?

* subject_age::subject_raceblack goes from .018 to .025 (1.4x increase) when limiting to discretionary searches

```{r logistic regression bin ages all searches}

logreg_oak3 <- logreg_oak1 %>%
  mutate(age3 = case_when(subject_age <= 24 ~ 0,
                          subject_age > 24 & subject_age <= 36 ~ 1,
                          subject_age > 36 ~ 2))

# about uniformly distributed... is that good? 1/3 of each stops are in each age cut off
ggplot(data = logreg_oak3) +
  geom_bar(mapping = aes(x = age3))

all_output3 <- glm(formula = search_conducted ~ age3*subject_race + factor(light_binary), data = logreg_oak3, family = binomial)

summary(all_output3)


```

```{r logistic regression BIN AGES and DISCRETIONARY SEARCHES only}

logreg_oak4 <- logreg_oak2 %>%
    mutate(age3 = case_when(subject_age <= 24 ~ 0,
                          subject_age > 24 & subject_age <= 36 ~ 1,
                          subject_age > 36 ~ 2))

all_output4 <- glm(formula = search_conducted ~ age3*subject_race + factor(light_binary), data = logreg_oak4, family = binomial)

summary(all_output4)

```


```{r}
#Each query returns an R dataframe
DBI::dbGetQuery(con, "SHOW TABLES")
```

```{r}
real.estate <- read.table("http://pages.pomona.edu/~jsh04747/courses/math154/CA_housedata.txt", 
                          header=TRUE)

logreg_oak2$search_conducted.f <- factor(logreg_oak2$search_conducted, labels = c("FALSE", "TRUE"))
logreg_oak2$search_conducted.f


set.seed(4747)
fitControl <- caret::trainControl(method="none")
tr.oak <- caret::train(search_conducted ~ subject_age + light_binary + subject_sex + lat_num + lng_num + subject_race, 
                      data = logreg_oak2,
                      na.action = na.exclude,
                      method="rpart2", 
                      trControl = fitControl, 
                      tuneGrid= data.frame(maxdepth=3))

rpart.plot::rpart.plot(tr.oak$finalModel)

tr.oak

tr.oak1 <- caret::train(search_conducted.f ~ subject_age + light_binary + subject_sex + lat_num + lng_num + subject_race, 
                      data = logreg_oak2,
                      na.action = na.exclude,
                      method="rpart2", 
                      trControl = fitControl, 
                      tuneGrid= data.frame(maxdepth=5))

rpart.plot::rpart.plot(tr.oak1$finalModel)

tr.oak1
```

```{r}

all_output_fix <- glm(formula = search_conducted ~ subject_age*subject_race + factor(light_binary) + arrest_made, data = logreg_oak1, family = binomial)

summary(all_output_fix)
as.data.frame(coef(all_output_fix))

```


```{r}
#coverage of subject age

CAoak %>%
  mutate(subject_age_recorded = case_when(is.na(subject_age) ~ 0,
                                          !is.na(subject_age) ~ 1)) %>%
  ggplot() + 
  geom_point(mapping = aes(x = nice_date, y = subject_age_recorded)) 
    
# only year 2017 and up
     
CAoak %>%
  mutate(subject_age_recorded = case_when(is.na(subject_age) ~ 0,
                                          !is.na(subject_age) ~ 1)) %>%
  filter(nice_year >= 2017) %>%
  ggplot() + 
  geom_point(mapping = aes(x = nice_date, y = subject_age_recorded), alpha =.4) 


is.na(CAoak$subject_age[2])

```